Reproducible Research Peer Assignment #1
========================================

## Loading and preprocessing the data
Activity.csv data will be read into R using read.csv
```{r results='asis'}
activity<-read.csv('/Users/Cherry/Downloads/activity.csv')
```

Take a peak at the data using dplyr
```{r , message=FALSE}
library(dplyr)
activity<-tbl_df(activity)
str(activity)
```

Date is currently a factor and will change it date class
```{r}
activity$datec<-as.character(activity$date)
activity$dated<-as.Date(activity$datec)
```

## What is mean total number of steps taken per day?
First, generate a histogram of total number of steps taken each day
```{r, fig.width=6, fig.height=6}
activity1<-
  activity %>%
  group_by(date) %>%
  summarize(totsteps=sum(steps, na.rm=TRUE))

hist(activity1$totsteps, main='Histogram of Total Number of Steps by Day', xlab='Total Number of Steps', ylab='Frequency', col='blue', breaks=10)
```

Find the mean and median of total number of steps taken per day
```{r}
mean(activity1$totsteps, na.rm=TRUE)
median(activity1$totsteps, na.rm=TRUE)
```

## What is the average daily activity pattern?
Via figure, explore the pattern of average steps taken across days over 5-minute intervals
```{r, fig.height=4, fig.width=13, message=FALSE}
activity1<-
  activity %>%
  group_by(interval) %>%
  summarise(avgsteps=mean(steps, na.rm=TRUE))

with(activity1, plot(interval, avgsteps, main='Time Series Plot of Average Daily Activity', xlab='5-minute interval', ylab='Average Steps Across Days', type='l'))
library(Hmisc)
minor.tick(nx=10)
```

Based on the figure, the highest average steps is around 850. Let's find out exactly when.
```{r}
filter(activity1, avgsteps==max(avgsteps))
```

## Imputing missing values
Obtain the number of missings in each variable
```{r}
sum(is.na(activity$date))
sum(is.na(activity$steps))
sum(is.na(activity$interval))
```

There are 2304 observations with missing steps, will replace them with mean of steps, save them in new dataset called activity.imp.
```{r}
activity.imp<-activity
activity.imp$steps[is.na(activity.imp$steps)]<-mean(activity.imp$steps, na.rm=TRUE)
sum(is.na(activity.imp$steps))
```

Regenrate histogram of total number of steps take each day, and calculate mean and median total number of steps taken per day.
```{r, fig.width=6, fig.height=6}
activity.imp1<-
  activity.imp %>%
  group_by(date) %>%
  summarise(totsteps=sum(steps, na.rm=TRUE)) 

hist(activity.imp1$totsteps, main='Histogram of Total Number of Steps by Day w/o Missing', xlab='Total Number of Steps', ylab='Frequency', col='blue', breaks=10)

mean(activity.imp1$totsteps, na.rm=TRUE)
median(activity.imp1$totsteps, na.rm=TRUE)
```
Mean and median of total number of steps taken in a day are different now after imputation.

## Are there differences in activity patterns between weekday and weekends?
Create variables for day of the week (day), and whether it's a weekday (weekday)
```{r}
activity.imp$day<-weekdays(activity$dated)
activity.imp$weekday<-ifelse(activity.imp$day %in% c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday'), 'Weekday', 'Weekend')
activity.imp$weekday<-as.factor(activity.imp$weekday)
```

A panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

First, create a summary dataset of average number of steps taken by weekday and interval uding dplyr.
```{r}
activity.imp1<-
  activity.imp %>%
  group_by(weekday, interval) %>%
  summarise(avgsteps=mean(steps))
```

Then apply ggplot2 to generate panel plot.
```{r, message=FALSE}
library(ggplot2)
ggplot(activity.imp1, aes(x=interval, y=avgsteps, colour=weekday)) + 
  geom_line() +
  xlab("Interval") +
  ylab("Average Number of Steps Taken") +
  ggtitle("Average Number of Steps Taken by Weekday or Weekend") +
  facet_wrap(~weekday)
```


